\documentclass[11pt,nocut]{article}

\usepackage{style/packages}
\usepackage{style/notations}
\usepackage{fancyhdr}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\fancyhead[R]{DS-GA 1014---Fall 2023}
\fancyhead[L]{HW9 --- Convexity --- due November 19, 2023 at noon}

\setcounter{section}{9}

\begin{document}
	
	\input{style/preamble_homeworks.tex}
	
	\vspace{5mm}

\begin{problem}[2 points]
	Decide if there exists a function $f: \R\to\R$ that satisfies the following properties. If there exists such a function, construct an example and prove that the properties are satisfied; if not, prove that there is a contradiction.
	\begin{enumerate}[label=\normalfont(\textbf{\alph*})]
		\item $f: \R\to\R$ is both convex and bounded, i.e., $\sup_{x\in\R} |f(x)| \le C$ for some constant $C$. 
		\item $f: \R\to\R$ is both strictly convex and bounded. 
	\end{enumerate}
\end{problem}

\vspace{5mm}


\begin{problem}[3 points]
	Let $A \in \R^{n \times m}$ and $y \in \R^n$.
	For $x \in \R^m$ we define
	$$
	f(x) = \|Ax-y\|^2.
	$$
	We will see a lot of this function when we discuss linear regression next lecture. Here we prove some useful properties of this function.
	\begin{enumerate}[label=\normalfont(\textbf{\alph*})]
		\item Compute the gradient $\nabla f(x)$ and the Hessian $H_f(x)$. 
            \item Compute the Hessian $H_f(x)$. 
		\item Show that $f$ is convex. 
	\end{enumerate}
\end{problem}


\vspace{5mm}

\begin{problem}[3 points] ~
\begin{enumerate}[label=\normalfont(\textbf{\alph*})]
\item Show that the function $f(x) = \log(x)$ is concave, that is, $-f(x)$ is convex. 
\item Use the concavity of $\log$, prove that for $p,q>1, \frac{1}{p} + \frac{1}{q} = 1$, and any non-negative reals $x,y\in\R_+$, we have the inequality
$$
xy \le \frac{x^p}{p} + \frac{x^q}{q}. 
$$
\underline{Hint:} evaluate the logarithm of the above expression.  
\item With the above result, show that for $p,q>1, \frac{1}{p} + \frac{1}{q} = 1$, and any $x,y\in\R^n$, we have
$$
\sum_{i=1}^n |x_i y_i| \le \|x\|_p\|y\|_q,  
$$
where $\|x\|_p = \left(\sum_{i=1}^n |x_i|^p\right)^{1/p}$ is the $\ell_p$ norm. \\
\underline{Hint:} apply the inequality in part $(b)$ to the product $\frac{|x_i|}{\|x\|_p}\frac{|y_i|}{\|y\|_q}$. 
\end{enumerate}
\end{problem}


\vspace{5mm}


\begin{problem}[2 points]
	Let $f: \R^n \to \R$ be a convex function. Assume that the minimum $m \defeq \min_{x \in \R^n} f(x)$ of $f$ on $\R^n$ is finite, and that the set of minimizers of $f$
	$$
	\mathcal{M} \defeq \big\{ v \in \R^n \, \big| \, f(v) = m \big\}
	$$
	is non-empty.
	
	\begin{enumerate}[label=\normalfont(\textbf{\alph*})]
		\item Show that $\mathcal{M}$ is a convex set.
		\item Show that if $f$ is strictly convex, then $\mathcal{M}$ has only one element.
	\end{enumerate}
\end{problem}


\vspace{5mm}


\begin{problem}[*]
Given positive definite matrix $A\in\R^{n\times n}$, consider the following non-negative cost function $f: \R^n\to\R$,
$$
f(x) = \|A - xx^\top\|_F^2, \quad x\in\R^n, 
$$
where the Frobenius norm of a matrix $M$ is defined as $\|M\|_F = \sqrt{\Tr(M^\top M)}$. \\
Prove or disprove: $f$ is convex. 
\end{problem}


\end{document}
