\documentclass[11pt,nocut]{article}

\usepackage{packages}
\usepackage{notations}
\usepackage{fancyhdr}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\fancyhead[R]{DS-GA 1014---Fall 2023}
\fancyhead[L]{HW7---Principal Component Analysis---due November 5, 2023 }


\setcounter{section}{7}

\begin{document}
% \maketitle

\input{../preamble_homeworks.tex}
\begin{problem}[3 points]
	We say that a symmetric matrix $M \in \R^{n \times n}$ is positive semi-definite if for all $x \in \R^n$, $
	x^{\sT} M x \geq 0$.
	\begin{enumerate}[label=\normalfont(\textbf{\alph*})]
  \item Let $D \in \R^{n \times n}$ be a diagonal matrix. When is $D$ positive semi-definite?
  \item Let $M \in \R^{n \times n}$ be a symmetric matrix. Show that $M$ is positive semi-definite \textbf{if and only if} its eigenvalues are all non-negative.
  \item Let $A \in \R^{n \times m}$ be any rectangular matrix. Show that $A^\top A$ and $AA^\top$ are positive semi-definite. (This shows that these matrices have non-negative eigenvalues.)
	\end{enumerate}
\end{problem}

\vspace{5mm}

\begin{problem}[3 points]
	Let $A \in \mathbb{R}^{n \times n}$ be a symmetric matrix.
    \begin{enumerate}[label=\normalfont(\textbf{\alph*})]
    \item Let $A = PDP^\top$ with $P$ orthogonal and $D$ diagonal. Show that for any $k \in \mathbb N$, $A^k = P D^k P^\top$.
    \item Same question when $k$ is a negative integer.
    \item Assume that A is positive semi-definite. Prove that there exists a symmetric positive semi-definite matrix $B \in \mathbb{R}^{n \times n}$ such that $A = B^2$. (Hint: in some sense, $B = A^{1/2}$. Can you guess how to define $B$?)
    \end{enumerate}
\end{problem}

\vspace{5mm}

\begin{problem}[2 points]
  Consider a dataset $x_1, \dots, x_n \in \R^d$ with mean $\mu \in \R^d$ nd covariance $\Sigma \in \R^{d \times d}$
	\begin{enumerate}[label=\normalfont(\textbf{\alph*})]
  \item Let $A \in \R^{m \times d}$ and $b \in \R^m$. Define $y_i = Ax_i + b$ for $i = 1,\dots, n$. Calculate the mean $\mu'$ and covariance $\Sigma'$ of the dataset $y_1, \dots, y_n$.
  \item Find $A$ and $b$ as a function of $\mu$ and $\Sigma$ so that $\mu' = 0$ and $\Sigma' = \mathrm{Id}$ (there are several solutions for $A$, you do not have to find all of them). We call such an operation ``whitening''. (Hint: search for $A$ of the form $PD'P^\top$, where $\Sigma = PDP^\top$ using the spectral theorem.)
  \end{enumerate}
\end{problem}


\vspace{5mm}

\begin{problem}[3 points]
  Complete the \texttt{mnist\_pca.ipynb} Jupyter notebook to compute the mean, covariance, and PCA of the MNIST dataset.
  Please only submit a pdf version of your notebook (right-click $\to$ `print' $\to$ `Save as pdf').
\end{problem}


\vspace{5mm}

\begin{problem}[$\star$]
  Let $A \in \R^{n \times m}$ a rectangular matrix. Show that there exists an orthonormal basis $u_1, \dots, u_m$ of $\R^m$ such that $Au_1, \dots, Au_m$ is an orthogonal family (its vectors are pairwise orthogonal but not necessarily of norm one). (Hint: apply the spectral theorem to $A^{\top} A$.)
\end{problem}


% \centerline{\pgfornament[width=7cm]{87}}

%\bibliographystyle{plain}
%\bibliography{./references.bib}
\end{document}
