\documentclass[11pt,nocut]{article}

\usepackage{packages}
\usepackage{notations}
\usepackage{fancyhdr}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\fancyhead[R]{DS-GA 1014 - Fall 2023}
\fancyhead[L]{HW4 - Norms and Inner Products - due October 8, 2021 at noon}

\setcounter{section}{4}

\begin{document}

\input{preamble_homeworks.tex}

\bigskip

\begin{problem}[2.5 points]
Given $x\in\R^n$, define 
$$
N(x) := \max\{|x_1|,|x_2|,...,|x_n|\}. 
$$
\begin{itemize}
    \item[(a)] Show that this defines a valid norm, i.e., verify that $N(\cdot)$ satisfies the three norm properties. 
    \item[(b)] Let $x = (10,0)$ and $y = (9,9)$. Using the norm $N(\cdot)$, which of $x$ and $y$ has bigger norm? Using the Euclidean norm, which of $x$ and $y$ has bigger norm?
    \item[(c)] Show that $N(x) > 1$ implies the Euclidean norm of $x$ is also greater than 1, i.e., $\|x\|_2 > 1$. 
\end{itemize}
    
\end{problem}

\vspace{1mm}

\begin{problem}[2 points] 
Decide whether each of the following functions $N(\cdot)$ is a valid norm. If yes, you do not have to justify your answer. If no, then provide an explicit counterexample and state which norm property is violated. 
\begin{itemize}
    \item [(a)] $x\in\R^2$. $N(x) := x^\top A x$, where $A = \begin{bmatrix}
        2 & 0 \\ 0 & 1
    \end{bmatrix}\in\R^{2\times 2}$. 
    \item[(b)]  $x\in\R^2$. $N(x) := x^\top A x$, where $A = \begin{bmatrix}
        2 & 0 \\ 0 & -1
    \end{bmatrix}\in\R^{2\times 2}$. 
    \item[(c)] $x\in\R^3$. $N(x) := 2|x_1| + \max\{|x_2|,|x_3|\}$. 
    \item[(d)] $x\in\R^n$. $N(x) := \left(\prod_{i=1}^n |x_i|\right)^{1/n} + \max_{i\in [n]}\{|x_i|\}$. 
\end{itemize}
\end{problem}

\vspace{1mm} 

\begin{problem}[1 point]
    Show that for any vector $x \in \R^n$, $$\displaystyle \left(\sum_{k=1}^n x_k \right)^2 \leq n \sum_{k = 1}^n x_k^2.$$ 
Hint: use Cauchy-Schwarz. 
\end{problem}

\vspace{1mm}

\begin{problem}[1.5 points]
Consider the set $V$ of random variables (on a probability space $\Omega$) with a finite second moment (i.e., for $X\in V$, $\E[X^2]<\infty$), with the inner product $\langle X,Y \rangle = \E[XY]$. Prove the Cauchy-Schwarz inequality: 
$$
|\E[XY]| \le \sqrt{\E[X^2]\E[Y^2]}.   
$$
Hint: given constants $a,b \ge 0$, evaluate the expression $\E[(aX-bY)^2]$. Then find appropriate values of $a,b$ to prove the statement. 
\end{problem}
    

\begin{problem}[3 points]
    Let $S$ be a subspace of $\R^n$ with $x \mapsto P_S(x)$ the orthogonal projector of all $x \in \R^n$. We will use the Euclidean dot product as our inner product $\langle x, y\rangle = \sum_{i=1}^n x_i y_i$. Show that for any $x \in \R^n$, 
    \begin{enumerate}[label=\normalfont(\textbf{\alph*})]
        \item $\langle x, y\rangle = \langle P_S(x), y\rangle$, for any vector $y$ in $S$. 
        \item $x - P_S(x)$ is orthogonal to $S$,
        \item $\lVert P_S(x) \rVert \leq \lVert x \rVert$, and the equality holds when $x\in S$.
    \end{enumerate}  
Hint: express $y$ in terms of the orthonormal basis of $S$: $(v_1, v_2, ..., v_k)$. 
\end{problem}


\vspace{1mm}

\begin{problem}[$\star$] 
Given $x\in\R^n$ and $p\in (1,\infty)$, define the $\ell_p$-norm as 
$$\|x\|_p := \left(\sum_{i=1}^n |x_i|^p\right)^{1/p}.$$
Prove that the $\ell_p$-norm satisfies the triangle inequality, that is, for any $x,y\in\R^n$, show that
$$
\|x + y \|_p \le \|x\|_p + \|y\|_p. 
$$
You may use the following inequality, known as Hölder's inequality, without proof: 
for any $x,y\in\R^n$, and constants $p,q\in\R_+$ satisfying $\frac{1}{p} + \frac{1}{q}=1$, we have
$$
\sum_{i=1}^n |x_i y_i| \le \|x\|_p\|y\|_q. 
$$
\end{problem}


\end{document}
